\texttt{MSnbase}, efficient and elegant R-based processing and visualisation of raw mass spectrometry data

### 2020-05-09 msnbase2.Rnw
mention what to do after boxcar
provide more general info and clarifications about the package

In this software note, we present version 2 of the \texttt{MSnbase}
\cite{Gatto:2012} software, available from the
Bioconductor~\cite{Huber:2015} project. \texttt{MSnbase}, like other
software such as Python-based {pyOpenMS}~\cite{Rost:2014},
spectrum\_utils~\cite{Bittremieux:2020} or
Pyteomics~\cite{Goloborodko:2013}, offers a platform that lies between
low level libraries and end-user software. \texttt{MSnbase} provides a
flexible R~\cite{R} command line environment for metabolomics and
proteomics mass spectrometry-based applications. It allows a detailed
step-by-step processing, analysis and exploration of the data and
development of novel computational mass spectrometry
methods~\cite{Stanstrup:2019}.

\section{Software functionality}

@@

In the following sections, we describe \texttt{MSnbase}'s ability to efficiently
manage very large mass spectrometry data and experiments and how it can be used
for data processing and visualisation. We will also illustrate how it makes use
of the forward-pipe operator (\texttt{\%>\%}) defined in the \texttt{magrittr}
package. This operator has proved useful to develop non-trivial analyses by
combining individual functions into easily readable pipelines.

@@

>>>>

In this software note, we present version 2 of the
\texttt{MSnbase}~\cite{Gatto:2012} software, available from the
Bioconductor~\cite{Huber:2015} project. The package, like other
software such as Python-based {pyOpenMS}~\cite{Rost:2014},
spectrum\_utils~\cite{Bittremieux:2020} or
Pyteomics~\cite{Goloborodko:2013}, offers a platform that lies between
low level libraries and end-user software. \texttt{MSnbase} provides a
flexible R~\cite{R} command-line environment for metabolomics and
proteomics mass spectrometry-based applications. It lays out a sound
infrastructure to work with raw mass spectrometry, quantitative and
proteomics identification data. The package enables manipulation of
these data (for example subsetting, filtering, or accessing specific
parts thereof), a detailed step-by-step processing (for example
smoothing and centroiding of raw data, or normalisation and imputation
of quantitative data), analysis and visualisation of these data and
the development of novel computational mass spectrometry
methods~\cite{Stanstrup:2019}. Here, we focus on the new developments
pertaining to raw mass spectrometry data handling and processing.

\section{Infrastructure for raw data}

@@

In the following sections, we describe \texttt{MSnbase}'s ability to
efficiently handle very large mass spectrometry data and experiments
and how it can be used for data processing and visualisation. We will
also illustrate how it makes use of the forward-pipe operator
(\texttt{\%>\%}) defined in the \texttt{magrittr} package. This
operator has proved useful to develop non-trivial analyses by
combining individual functions into easily readable pipelines.

@@

After processing of the BoxCar data, the final object can either be
further analysed in R or written back to disk as an \texttt{mzML} file
using \texttt{writeMSData()} for processing with other tools.
